\center \fontsize{18.8}{20} \selectfont Towards Globally Color-Dual All-Loop Integrands

### 2023-09-21 notes/2loopNLSM.tex
changes over zoom

\usetikzlibrary{external}  
\immediate\write18{mkdir -p images}  %% Create `pgf-img` directory
\tikzexternalize[                     %% Activate externalization
prefix=images/,                    %% Avoid cluttering the directory
system call={                       %% Use lualatex in system call
   pdflatex \tikzexternalcheckshellescape -halt-on-error -shell-escape -interaction=batchmode -jobname="\image" "\texsource" || rm "\image.pdf"
},
]
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\emailAdd{alexander.edison@northwestern.edu, james.mangan@northwestern.edu, pavao@u.northwestern.edu}

@@

TODO:\\
For now, let's put the things we need to check/compute here in the abstract. 
\begin{enumerate}
\item Settle on $n$ vs $N$ vs $\mathcal{N}$ as the symbol for
  numerator
\item \ace{I am starting to get a bit concerned about the overuse of the
  phrase ``color-dual''}
\item \ace{I have defined a \textcolor{ucp-color}{\texttt{ucp-color}}
    , \textcolor{cut-color}{\texttt{cut-color}},
    \textcolor{refl-color}{\texttt{refl-color}}, and
    \textcolor{jacobi-color}{\texttt{jacobi-color}} as aliases for
    other colors so that we don't have to keep changing colors in
    multiple locations when we want to change diagram color schemes.
    I think I have updated all of the diagrams to use the
    correctly-named color, but haven't changed any of the backing
    colors for the names.}
\end{enumerate}
}

@@

Color-kinematics duality is a statement about the off-shell graphical
simplicity that manifest redundancy in the on-shell observables of
gauge and gravity theories. Ignoring factors of $i$ and the coupling
constant, a generic $m$-point $L$-loop amplitude in a color-dual
theory -- such as NLSM, YM, or bi-adjoint scalar
theory\ace{technically this is true as written for any theory.  In
  fact we slip that type of disclaimer in quite a lot throughout the
  rest of the section} \jm{I removed the second time we say `YM, NLSM,
  $\phi^3$' which was redundant.  I don't mind if we say it here
  because I want people to tell people that these are the kinds of
  theories we're thinking about.  We could omit it but then we'd have
  to say that the theory has at least one color factor and is probably
  an adjoint type theory and at that point I'd rather just give an
  example.} -- can be written as
\begin{equation}
  \mathcal{A}_m^L = \sum \limits_\Gamma \frac{1}{S_\Gamma} \int \frac{d^{LD}\ell}{(2\pi)^{LD}} \frac{c_\Gamma n_\Gamma}{d_\Gamma} .
  \label{eq:gen-amp}
\end{equation}
Here the spacetime dimension is $D$.  Even though the underlying
theory may have quartic (or higher multiplicity) vertices, it is
possible to multiply and divide by propagators so that the sum runs
over the set of purely cubic graphs, $\Gamma$.  Overcounting is removed by
the internal symmetry factor $S_\Gamma$, which tracks the number of
graph symmetries with external legs held fixed.  The color factor
$c_\Gamma$ is the string of structure constants $f^{abc}$ associated
with the graph $\Gamma$,\footnote{This paper is focused on theories
  with particles that transform in the adjoint.  However, the double
  copy can be applied to a far more diverse set of theories with
  particles that transform in the fundamental or more diverse
  representations, see for example \cite{Johansson:2017srf,
    Johansson2014zca, Johansson:2015oia, Johansson:2019dnu, Carrasco:2020ywq,
    Carrasco:2023vjg}.}  the denominator $d_\Gamma$
is the product of propagators associated with the graph $\Gamma$, and
the numerator $n_\Gamma$ is all of the remaining kinematic factors
including dot products of momenta and polarization vectors.

@@

\jm{The next section feels a little redundant and silly sometimes
  because the integrand, color factors, kinematic numerators,
  etc. were defined a while ago.  I'm not sure how to merge the two
  sections though.}

@@

\ace{Awkward phrasing/transition.}  What has just been described is
sometimes referred to as ``global'' color-kinematics duality because
the Jacobi identities are solved by a single (global) integrand with
off-shell internal lines.  In order to perform the double copy it is
enough to simply satisfy the Jacobi identities on the generalized
unitarity cuts, meaning that each graph can be assigned its own ansatz
\cite{Bern:2015ooa}.  If a global solution cannot be found, this is
one way to enlarge the ansatz where the other option described above
was to incorporate non-local terms in the ``numerators''.  A globally
color-dual integrand is desirable because of its simplicity and
because a color-dual Lagrangian would naively produce precisely this
object.

@@

Every numerator is required to respect all of the symmetries (or
automorphisms) of its associated graph.  Graph symmetries are an
avatar of Bose symmetry.  For a pure ansatz construction it is very
reasonable to impose graph symmetries, but color-kinematics duality at
tree-level is often achieved at the expense of manifest Bose symmetry
\cite{Cheung:2016prv, Cheung:2021zvb, Brandhuber:2021bsf,
  Edison:2020ehu}.  In an ansatz construction graph symmetries are
imposed for \emph{every} graph, not just the basis graphs.
%For example, the symmetries of the crossed box are spanned by
%\begin{equation}
%\CrossedBoxGraphSym{1}{2}{3}{4} \sim \CrossedBoxGraphSym{2}{1}{3}{4} \text{ and } \CrossedBoxGraphSym{1}{2}{3}{4} \sim \CrossedBoxGraphSym{2}{1}{4}{3}
% \end{equation}
\ace{is there a reason we commented out this diagram?} \jm{You can try
  either one.  I liked the one I left in better.}  For example, the
symmetries of the crossed box
\begin{equation}
\CrossedBoxGraphSym{1}{2}{3}{4}
\end{equation}
are spanned by $\{1234\} \to \{2134\}$ and $\{1234\} \to \{1243\}$.
In general, internal edge labels must be tracked as well.
Furthermore, care must be taken with the signs in a symmetry relation.
To compensate for the antisymmetric color factors, $f^{abc}$, graph
vertices are all totally antisymmetric and hence have signs built into
their orientations.  Once the signs are included in the symmetry
relations, the crossed box can be expressed in terms of the double box
and penta-triangle, thus imposing non-trivial constraints on the
ansatz. \ace{I'm not quite sure why the signs are mentioned in the
  previous sentence?}

@@

The final constraints comes from ensuring that the ansatz correctly
reproduces the generalized unitarity cuts of the desired theory, thus
guaranteeing that the numerators encode the correct physical,
theory-dependent information.  Generalized unitarity methods has been
reviewed in many places so we will be light on details here
\cite{BCFUnitarity, BernHuangReview, Bern:2015ooa, Bourjaily:2017wjl,
  JJHenrikReview, FivePointN4BCJ, Bern:2012uf, BCJreview,
  Edison:2022jln, Edison:2022smn}.  \ace{Here is a semi-random
  collection of unitarity-related citations.  Probably want to bulk it
  up a bit more.}  Recently, generalized unitarity was optimized for
even-multiplicity theories like NLSM \cite{Carrasco:2023qgz}.

@@

\section{One-loop cubic construction}
\label{sec:cubic}
Here we will describe why constructing the full Yang-Mills kinematic
algebra (outside the self-dual sector) is generically hard, and why
the same reason makes NLSM beyond one-loop hard as well. The major
obstacle stems from the vector state sum mixing with higher-point
contacts. Starting with the Yang-Mills Lagrangian,
\begin{equation}
  \label{eq:ym-lag}
  \mathcal{L}^{\text{YM}} = -\frac{1}{2}(\partial_\mu A_\nu)^2
  + f^{abc} \partial_\mu A^a_\nu A^b_\mu A^c_\nu
  + f^{abe}f^{ecd}A^a_\mu A^b_\nu  A^c_\mu A^d_\nu 
\end{equation}
we can express the YM three-point interaction in terms of the
following vertex in Lorenz gauge:
\begin{equation}
\Acubic{hgrey0}{}{nhpRed}{}{nhpRed}{fermion2}{} =(\varepsilon_i\varepsilon_j)(\varepsilon_k p_i).
\end{equation}
This kinematic vertex is anti-symmetric in $1\leftrightarrow 2$
exchange, and from it we can construct the full Lorenz gauge
Yang-Mills vertex by summing over cyclic permutations,
\begin{equation}
V^{(123)}_{\text{YM}} = \Acubic{hgrey0}{}{nhpRed}{}{nhpRed}{fermion2}{}  +\text{cyc}(123)\,.
\end{equation}
Contracting this with the vector state sum, we can see that the cubic
Yang-Mills vertex does not satisfy the four-point Jacobi identity:
\begin{equation}
  \label{cubicJac}
  {}_1\langle V^{2}_{\text{YM}}V^{3}_{\text{YM}}\rangle_4+\text{cyc}(234) = \varepsilon_{(12)}  \varepsilon_{(34)} (s_{13}-s_{23}) +\text{cyc}(234) \neq 0
\end{equation}
where $ \varepsilon_{(ij)} = \varepsilon_{i}\cdot \varepsilon_{j}$.
However, the remaining terms can be absorbed into the definition of the
four-point kinematic numerator, as follows
\begin{equation}
{}_1\langle V^{2}_{\text{YM}+B}V^{3}_{\text{YM}+B}\rangle_4 = {}_1\langle V^{2}_{\text{YM}}V^{3}_{\text{YM}}\rangle_4 + s_{12} (\varepsilon_{(13)}\varepsilon_{(24)}-\varepsilon_{(14)}\varepsilon_{(23)})
\end{equation}
Note that this new definition of the $s$-channel numerator contains two factors of $\varepsilon_{(ij)} $ that contract spacetime indices \textit{across} the factorization, suggesting the inclusion of a spin-2 mode. As such, we use $V_{\text{YM}+B}$ to denote the inclusion of an additional two-form, as was studied in Ref.~\cite{Ben-Shahar:2022ixa} for constructing the NMHV Yang-Mills Lagrangian.  After the introduction of the two-form, the
numerators now satisfy the Jacobi identity:
\begin{equation}
  {}_1\langle
  V^{2}_{\text{YM}+B}V^{3}_{\text{YM}+B}\rangle_4+\text{cyc}(234) =0\,,
\end{equation}
where the new cubic Lagrangian takes the form\,,
\begin{equation}
\mathcal{L}^{\text{YM}+B} = -\frac{1}{2}(\partial_\mu A_\nu)^2 -
B_{\mu\nu}\Box\tilde{B}_{\mu\nu}+ f^{abc} \partial_\mu A^a_\nu A^b_\mu
A^c_\nu+ f^{abc} (B_{\mu\nu}+\Box \tilde{B}_{\mu\nu})^a A^b_\mu
A^c_\nu\, .
\end{equation}
Of course, higher multiplicity would likely require further redefinition of
the three-point vertex to satisfy Jacobi identities on all internal
edges. While introducing successively higher spin states could in
principle work at tree level, we will see that it is not consistent
with what we find at general loop order. We will comment on this in
\cref{sec:Discussion} \ace{Is this the sec you are thinking of?} \nhp{ Yes. thanks :-)}. For
now, we will study how to construct one-loop integrands consistent
with the color-kinematics by isolating the cubic sector of the state
sum above.

As can be seen above, the Jacobi identity of \cref{cubicJac} fails
only in terms with two factors of polarization dot products,
$(\varepsilon\varepsilon)^2$. Whereas if we restricted ourselves to
just considering factors with a single polarization dot product, the
Jacobi identity of Yang-Mills would be satisfied \textit{off-shell}
and in arbitrary spacetime dimensions
\begin{equation}
{}_1\langle V^{2}_{\text{YM}}V^{3}_{\text{YM}}\rangle_4\big|^{(\varepsilon\varepsilon)^1}+\text{cyc}(234) =0\,.
\end{equation}
\ace{Should we mention that this is closely related to MHV vs NMHV?} \nhp{Yes, I added a couple sentences and reference below -- let me know what you think}.

This restriction to the $(\varepsilon\varepsilon)^1$ cubic sector of Yang-Mills is closely related to the MHV decomposition of color-dual numerators \cite{ElvangHuangReview,Chen:2019ywi,Chen:2021chy}, where choosing appropriate reference vectors sends all $(\varepsilon\varepsilon) \rightarrow 0$, with the exception of one of the positive helicity polarizations, $\varepsilon^+_1$, dotted into all the $n-2$ negative helicity modes in the full MHV amplitude, $\varepsilon^+_1 \varepsilon^-_i \neq 0$. We describe this in detail in \cref{Conventions}. With the interest in constructing dimension agnostic color-dual integrands at one-loop, we want to avoid making specific reference to 4D helicity states.

@@

\subsection{One-loop color-dual integrands}\label{oneLoopCK}
Our first application of this theory for color-dual construction at
loop level is for self-dual Yang-Mills (SDYM). At tree-level, the
off-shell cubic vertices can be written in terms of light cone
coordinates \cite{Monteiro2011pc}:
\begin{equation}
X(p,k) = p_u k_w-p_w k_u
\end{equation}
At loop-level, this construction needs to be analytically continued to
general dimension in order to apply dimensional regularization at
one-loop. This can be done using the cubic semi-YM vertices of the
previous section:
\begin{equation} \label{DdimSDYM}
\cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{}= \mathcal{X}(k_1,k_2) = (\varepsilon_1\bar{\varepsilon}_3)(\varepsilon_2 k_3) -  (\varepsilon_2\bar{\varepsilon}_3)(\varepsilon_1 k_3)
\end{equation}
Plugging in on-shell all-plus helicity states in light-cone gauge will
yield precisely the 4D SDYM vertex, up to an unphysical phase \ace{Can
  we provide an example?} \nhp{yes, provided}:
  \be
  \mathcal{X}(k^+_1,k^+_2) = \frac{\langle 12\rangle^3}{ \langle23\rangle \langle 31\rangle} \sim \langle 12\rangle = k_{1,u} k_{2,w}-k_{1,w} k_{2,u}
 \ee
where $\mathcal{X}(k^+_1,k^+_2) = A^{\text{YM}}(1^+,2^+,3^-)$ in light cone gauge, and in the second step applied momentum conservation to redefinition of the spinor bracket, $\langle 12\rangle \rightarrow X(p,k)$ defined in \cref{Conventions}. Of course, the form of \cref{DdimSDYM} has the advantage of
permitting a $D$-dimensional construction of the one-loop
integrand. Below we provide a four-point example:
\begin{equation}
\simplebox = g_{\text{YM}}^4 \langle \mathcal{X}(k_1,l_1)\mathcal{X}(k_2,l_2)\mathcal{X}(k_3,l_3)\mathcal{X}(k_4,l_4)\rangle
\end{equation}
where $l_i = l+k_1+k_2+\cdots+ k_i$ and the bracket $\langle \cdots\rangle$ indicates that we have applied the off-shell state projector, $\sum {\varepsilon^\mu_{(+l_i)}\varepsilon^\nu_{(-l_i)}}=P_{(l_i)}^{\mu\nu}$ on all internal polarizations. In general, the $n$-gon vertex would be
\begin{equation}
\mathcal{N}^{\text{SDYM}}_{n\text{-gon}}= \nGonVector = g_{\text{YM}}^n \langle \mathcal{X}(k_1,l_1)\mathcal{X}(k_2,l_2)\cdots \mathcal{X}(k_n,l_n) \rangle
\end{equation}
Due to the state-sum of internal loop factors,
$\sum {\varepsilon_{(+l)}\varepsilon_{(-l)}}\sim D$, the integrand above
depends explicitly on the spacetime dimension, $D$. By taking a
derivative\footnote{Another way to understand this derivative is that
  it selects the large $D$ behavior of the integrand.}, we can recover
the integrand needed for all-plus one-loop amplitudes with an internal
scalar:
\begin{equation}
\partial_D \mathcal{N}^{\text{SDYM}}_{n\text{-gon}}=\nGonScalar  = (2 g_{\text{YM}})^n (l_1 \varepsilon_1)(l_2 \varepsilon_2)\cdots (l_n \varepsilon_n)
\end{equation}
This integrand numerator with internal scalar loop is precisely what
one would obtain from the Feynman rules of YZ-theory, absent the $\bar{Z}Z$
internal vector loop. For more background, we refer the reader to
\cref{sec:CKLagrangians}. When plugging on the on-shell states of $YZ$-theory, we thus obtain the following expression for the NLSM one-loop $n$-gon numerator:
\begin{equation}
 \mathcal{N}^{\text{NLSM}}_{n\text{-gon}}=\left[\partial_D  \mathcal{N}^{\text{SDYM}}_{n\text{-gon}}\right]^{\epsilon \rightarrow k}=\nGonScalar =  f_\pi^{-n} [12][23]\cdots [n1]
\end{equation}
Where we have define the antisymmetric kinematic variable, $[LR] = l_L^2 - l_R^2$, since when plugging on longitudinal modes, the $n$-gon can be re-expressed completely in terms of inverse propagators, $2(k_i \cdot l_i) = l_{i}^2-l^2_{i-1}$. We have verified through 10-point one-loop that this $n$-gon expression is a valid color-dual representation for NLSM.

>>>>

%\usetikzlibrary{external}  
%\immediate\write18{mkdir -p images}  %% Create `pgf-img` directory
%\tikzexternalize[                     %% Activate externalization
%prefix=images/,                    %% Avoid cluttering the directory
%system call={                       %% Use lualatex in system call
  % pdflatex \tikzexternalcheckshellescape -halt-on-error -shell-escape -interaction=batchmode -jobname="\image" %"\texsource" || rm "\image.pdf"
%},
%]
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\emailAdd{alexander.edison@northwestern.edu, 
\\
james.mangan@northwestern.edu, 
\\
pavao@u.northwestern.edu}

@@

TODO:\\
For now, let's put the things we need to check/compute here in the abstract. 
\begin{enumerate}
%\item Settle on $n$ vs $N$ vs $\mathcal{N}$ as the symbol for
 % numerator

\item \ace{I am starting to get a bit concerned about the overuse of the
  phrase ``color-dual''} \nhp{We all choose capital $N$}
\item \ace{I have defined a \textcolor{ucp-color}{\texttt{ucp-color}}
    , \textcolor{cut-color}{\texttt{cut-color}},
    \textcolor{refl-color}{\texttt{refl-color}}, and
    \textcolor{jacobi-color}{\texttt{jacobi-color}} as aliases for
    other colors so that we don't have to keep changing colors in
    multiple locations when we want to change diagram color schemes.
    I think I have updated all of the diagrams to use the
    correctly-named color, but haven't changed any of the backing
    colors for the names.}
\end{enumerate}
}

@@

Color-kinematics duality is a statement about the off-shell graphical
simplicity that manifest redundancy in the on-shell observables of
gauge and gravity theories. Ignoring factors of $i$ and the coupling
constant, a generic $m$-point $L$-loop amplitude in a color-dual
theory -- such as NLSM, YM, or bi-adjoint scalar
theory\ace{technically this is true as written for any theory.  In
  fact we slip that type of disclaimer in quite a lot throughout the
  rest of the section} \jm{I removed the second time we say `YM, NLSM,
  $\phi^3$' which was redundant.  I don't mind if we say it here
  because I want people to tell people that these are the kinds of
  theories we're thinking about.  We could omit it but then we'd have
  to say that the theory has at least one color factor and is probably
  an adjoint type theory and at that point I'd rather just give an
  example.} \nhp{Nic will clarify between cubic vs color-dual here}-- can be written as
\begin{equation}
  \mathcal{A}_m^L = \sum \limits_\Gamma \frac{1}{S_\Gamma} \int \frac{d^{LD}\ell}{(2\pi)^{LD}} \frac{c_\Gamma n_\Gamma}{d_\Gamma} .
  \label{eq:gen-amp}
\end{equation}
Here the spacetime dimension is $D$.  Even though the underlying
theory may have quartic (or higher multiplicity) vertices, it is
possible to multiply and divide by propagators so that the sum runs
over the set of purely cubic graphs, $\Gamma$.  Overcounting is removed by
the internal symmetry factor $S_\Gamma$, which tracks the number of
graph symmetries with external legs held fixed.  The color factor
$c_\Gamma$ is the string of structure constants $f^{abc}$ associated
with the graph $\Gamma$,\footnote{This paper is focused on theories
  with particles that transform in the adjoint.  However, the double
  copy can be applied to a far more diverse set of theories with
  particles that transform in the fundamental or more diverse
  representations, see for example \cite{Johansson:2017srf,
    Johansson2014zca, Johansson:2015oia, Johansson:2019dnu, Carrasco:2020ywq,
    Carrasco:2023vjg}.}  the denominator $d_\Gamma$
is the product of propagators associated with the graph $\Gamma$, and
the numerator $n_\Gamma$ is all of the remaining kinematic factors
including dot products of momenta and polarization vectors.

@@

@@

\ace{Awkward phrasing/transition.} \nhp{will message language} What has just been described is
sometimes referred to as ``global'' color-kinematics duality because
the Jacobi identities are solved by a single (global) integrand with
off-shell internal lines.  In order to perform the double copy it is
enough to simply satisfy the Jacobi identities on the generalized
unitarity cuts, meaning that each graph can be assigned its own ansatz
\cite{Bern:2015ooa}.  If a global solution cannot be found, this is
one way to enlarge the ansatz where the other option described above
was to incorporate non-local terms in the ``numerators''.  A globally
color-dual integrand is desirable because of its simplicity and
because a color-dual Lagrangian would naively produce precisely this
object.

@@

Every numerator is required to respect all of the symmetries (or
automorphisms) of its associated graph.  Graph symmetries are an
avatar of Bose symmetry.  For a pure ansatz construction it is very
reasonable to impose graph symmetries, but color-kinematics duality at
tree-level is often achieved at the expense of manifest Bose symmetry
\cite{Cheung:2016prv, Cheung:2021zvb, Brandhuber:2021bsf,
  Edison:2020ehu}.  In an ansatz construction graph symmetries are
imposed for \emph{every} graph, not just the basis graphs.
For example, the symmetries of the crossed box are spanned by,
\nhp{Alex, fix the scaling to get the boxes to fit}
\be
\CrossedBoxGraphSym{1}{2}{3}{4}\!\! \sim\!\! \CrossedBoxGraphSym{2}{1}{3}{4} 
\text{ and }
\CrossedBoxGraphSym{1}{2}{3}{4} \!\!\sim\!\! \CrossedBoxGraphSym{2}{1}{4}{3}
\ee
%\ace{is there a reason we commented out this diagram?} \jm{You can try
 % either one.  I liked the one I left in better.}  For example, the
%symmetries of the crossed box
%\begin{equation}
%\CrossedBoxGraphSym{1}{2}{3}{4}
%\end{equation}
which are just functional relabelings of external legs, $\{1234\} \to \{2134\}$ and $\{1234\} \to \{1243\}$.
In general, internal edge labels must be tracked as well.
Furthermore, care must be taken with the signs in a symmetry relation.
To compensate for the antisymmetric color factors, $f^{abc}$, graph
vertices are all totally antisymmetric and hence have signs built into
their orientations.  Once imposing symmetry relations, the crossed box can be expressed in terms of the double box
and penta-triangle, thus imposing non-trivial constraints on the
ansatz. 
\nhp{Someone add in comment here about how symmetry relations imposes constraints of jacobi from equation above}.
%\ace{I'm not quite sure why the signs are mentioned in the
%  previous sentence?}

@@

The final constraints comes from ensuring that the ansatz correctly
reproduces the generalized unitarity cuts of the desired theory, thus
guaranteeing that the numerators encode the correct physical,
theory-dependent information.  Generalized unitarity methods has been
reviewed in many places so we will be light on details here
\cite{BCFUnitarity, Bern:2015ooa, Bourjaily:2017wjl,
  JJHenrikReview, FivePointN4BCJ, Bern:2012uf, BCJreview,
  Edison:2022jln, Edison:2022smn}. Recently, generalized unitarity was optimized for
even-multiplicity theories like NLSM \cite{Carrasco:2023qgz}.

@@

\section{One-loop cubic construction}
\label{sec:cubic}
Here we will describe why constructing the full Yang-Mills kinematic
algebra (outside the self-dual sector) is generically hard, and why
the same reason makes NLSM beyond one-loop hard as well. The major
obstacle stems from the vector state sum mixing with higher-point
contacts. Starting with the Yang-Mills Lagrangian,
\begin{equation}
  \label{eq:ym-lag}
  \mathcal{L}^{\text{YM}} = -\frac{1}{2}(\partial_\mu A_\nu)^2
  + f^{abc} \partial_\mu A^a_\nu A^b_\mu A^c_\nu
  + f^{abe}f^{ecd}A^a_\mu A^b_\nu  A^c_\mu A^d_\nu 
\end{equation}
we can express the YM three-point interaction in terms of the
following vertex in Lorenz gauge:
\begin{equation}
\Acubic{hgrey0}{}{nhpRed}{}{nhpRed}{fermion2}{} =(\varepsilon_i\varepsilon_j)(\varepsilon_k p_i).
\end{equation}
This kinematic vertex is anti-symmetric in $1\leftrightarrow 2$
exchange, and from it we can construct the full Lorenz gauge
Yang-Mills vertex by summing over cyclic permutations,
\begin{equation}
V^{(123)}_{\text{YM}} = \Acubic{hgrey0}{}{nhpRed}{}{nhpRed}{fermion2}{}  +\text{cyc}(123)\,.
\end{equation}
Contracting this with the vector state sum, we can see that the cubic
Yang-Mills vertex does not satisfy the four-point Jacobi identity:
\begin{equation}
  \label{cubicJac}
  {}_1\langle V^{2}_{\text{YM}}V^{3}_{\text{YM}}\rangle_4+\text{cyc}(234) = \varepsilon_{(12)}  \varepsilon_{(34)} (s_{13}-s_{23}) +\text{cyc}(234) \neq 0
\end{equation}
where $ \varepsilon_{(ij)} = \varepsilon_{i}\cdot \varepsilon_{j}$.
However, the remaining terms can be absorbed into the definition of the
four-point kinematic numerator, as follows
\begin{equation}
{}_1\langle V^{2}_{\text{YM}+B}V^{3}_{\text{YM}+B}\rangle_4 = {}_1\langle V^{2}_{\text{YM}}V^{3}_{\text{YM}}\rangle_4 + s_{12} (\varepsilon_{(13)}\varepsilon_{(24)}-\varepsilon_{(14)}\varepsilon_{(23)})
\end{equation}
Note that this new definition of the $s$-channel numerator contains two factors of $\varepsilon_{(ij)} $ that contract spacetime indices \textit{across} the factorization, suggesting the inclusion of a spin-2 mode. As such, we use $V_{\text{YM}+B}$ to denote the inclusion of an additional two-form, as was studied in Ref.~\cite{Ben-Shahar:2022ixa} for constructing the NMHV Yang-Mills Lagrangian.  After the introduction of the two-form, the
numerators now satisfy the Jacobi identity:
\begin{equation}
  {}_1\langle
  V^{2}_{\text{YM}+B}V^{3}_{\text{YM}+B}\rangle_4+\text{cyc}(234) =0\,,
\end{equation}
where the new cubic Lagrangian takes the form\,,
\begin{equation}
\mathcal{L}^{\text{YM}+B} = -\frac{1}{2}(\partial_\mu A_\nu)^2 -
B_{\mu\nu}\Box\tilde{B}_{\mu\nu}+ f^{abc} \partial_\mu A^a_\nu A^b_\mu
A^c_\nu+ f^{abc} (B_{\mu\nu}+\Box \tilde{B}_{\mu\nu})^a A^b_\mu
A^c_\nu\, .
\end{equation}
Of course, higher multiplicity would likely require further redefinition of
the three-point vertex to satisfy Jacobi identities on all internal
edges. While introducing successively higher spin states could in
principle work at tree level, we will see that it is not consistent
with what we find at general loop order. We will comment on this in
\cref{sec:Discussion}. For
now, we will study how to construct one-loop integrands consistent
with the color-kinematics by isolating the cubic sector of the state
sum above.

As can be seen above, the Jacobi identity of \cref{cubicJac} fails
only in terms with two factors of polarization dot products,
$(\varepsilon\varepsilon)^2$. Whereas if we restricted ourselves to
just considering factors with a single polarization dot product, the
Jacobi identity of Yang-Mills would be satisfied \textit{off-shell}
and in arbitrary spacetime dimensions
\begin{equation}
{}_1\langle V^{2}_{\text{YM}}V^{3}_{\text{YM}}\rangle_4\big|^{(\varepsilon\varepsilon)^1}+\text{cyc}(234) =0\,.
\end{equation}
This restriction to the $(\varepsilon\varepsilon)^1$ cubic sector of Yang-Mills is closely related to the MHV decomposition of color-dual numerators \cite{TasiLance,ElvangHuangReview,Chen:2019ywi,Chen:2021chy}, where choosing appropriate reference vectors sends all $(\varepsilon\varepsilon) \rightarrow 0$, with the exception of one of the positive helicity polarizations, $\varepsilon^+_1$, dotted into all the $n-2$ negative helicity modes in the full MHV amplitude, $\varepsilon^+_1 \varepsilon^-_i \neq 0$. We describe this in detail in \cref{Conventions}. With the interest in constructing dimension agnostic color-dual integrands at one-loop, we want to avoid making specific reference to 4D helicity states.

@@

\subsection{One-loop color-dual integrands}\label{oneLoopCK}
Our first application of this theory for color-dual construction at
loop level is for self-dual Yang-Mills (SDYM). At tree-level, the
off-shell cubic vertices can be written in terms of light cone
coordinates \cite{Monteiro2011pc}:
\begin{equation}
X(p,k) = p_u k_w-p_w k_u
\end{equation}
At loop-level, this construction needs to be analytically continued to
general dimension in order to apply dimensional regularization at
one-loop. This can be done using the cubic semi-YM vertices of the
previous section:
\begin{equation} \label{DdimSDYM}
\cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{}= \mathcal{X}(k_1,k_2) = (\varepsilon_1\bar{\varepsilon}_3)(\varepsilon_2 k_3) -  (\varepsilon_2\bar{\varepsilon}_3)(\varepsilon_1 k_3)
\end{equation}
Plugging in on-shell all-plus helicity states in light-cone gauge will
yield precisely the 4D SDYM vertex, up to an unphysical phase \ace{Can
  we provide an example?} \nhp{yes, provided}:
  \be
  \mathcal{X}(k^+_1,k^+_2) = \frac{\langle 12\rangle^3}{ \langle23\rangle \langle 31\rangle} \sim \langle 12\rangle = k_{1,u} k_{2,w}-k_{1,w} k_{2,u}
 \ee
where $\mathcal{X}(k^+_1,k^+_2) = A^{\text{YM}}(1^+,2^+,3^-)$ in light cone gauge, and in the second step applied momentum conservation to redefinition of the spinor bracket, $\langle 12\rangle \rightarrow X(p,k)$ defined in \cref{Conventions}. Of course, the form of \cref{DdimSDYM} has the advantage of
permitting a $D$-dimensional construction of the one-loop
integrand. Below we provide a four-point example: \nhp{change ordering on legs}
\begin{equation}
\simplebox = g_{\text{YM}}^4 \langle \mathcal{X}(k_1,l_1)\mathcal{X}(k_2,l_2)\mathcal{X}(k_3,l_3)\mathcal{X}(k_4,l_4)\rangle
\end{equation}
where $l_i = l+k_1+k_2+\cdots+ k_i$ and the bracket $\langle \,\cdots\rangle$ indicates that we have applied the gauge fixed state projector, $\sum {\varepsilon^\mu_{(+l_i)}\varepsilon^\nu_{(-l_i)}}=\eta^{\mu\nu}$ on all internal polarizations. In general, the $n$-gon diagram would be,
\begin{equation}
\mathcal{N}^{\text{SDYM}}_{n\text{-gon}}= \nGonVector = g_{\text{YM}}^n \langle \mathcal{X}(k_1,l_1)\mathcal{X}(k_2,l_2)\cdots \mathcal{X}(k_n,l_n) \rangle
\end{equation}
Due to the state-sum of internal loop factors,
$\sum {\varepsilon_{(+l)}\varepsilon_{(-l)}}\sim D$, the integrand above
depends explicitly on the spacetime dimension, $D$. By taking a
derivative\footnote{Another way to understand this derivative is that
  it selects the large $D$ behavior of the integrand.}, we can recover
the integrand needed for all-plus one-loop amplitudes with an internal
scalar:
\begin{equation}
\partial_D \mathcal{N}^{\text{SDYM}}_{n\text{-gon}}=\nGonScalar  = (2 g_{\text{YM}})^n (l_1 \varepsilon_1)(l_2 \varepsilon_2)\cdots (l_n \varepsilon_n)
\end{equation}
This integrand numerator with internal scalar loop is precisely what
one would obtain from the Feynman rules of YZ-theory, absent the $\bar{Z}Z$
internal vector loop. For more background, we refer the reader to
\cref{sec:CKLagrangians}. When plugging on the on-shell states of $YZ$-theory, we thus obtain the following expression for the NLSM one-loop $n$-gon numerator:
\begin{equation}
 \mathcal{N}^{\text{NLSM}}_{n\text{-gon}}=\left[\partial_D  \mathcal{N}^{\text{SDYM}}_{n\text{-gon}}\right]^{\epsilon \rightarrow k}=\nGonScalar =  f_\pi^{-n} [12][23]\cdots [n1]
\end{equation}
Where we have define the antisymmetric kinematic variable, $[ij] = l_i^2 - l_j^2$, since when plugging on longitudinal modes, the $n$-gon can be re-expressed completely in terms of inverse propagators, $2(k_i \cdot l_i) = l_{i}^2-l^2_{i-1}$. We have verified through 10-point one-loop that this $n$-gon expression is a valid color-dual representation for NLSM.

Before proceeding, we note that the above definition does give rise to ``pathological" bubble-on-external-leg (BEL) diagrams, discussed previously in \cref{sec:bootstrap}. However, one can show that these diagrams integrate to zero for spacetime dimension, $D>2$, and thus can be disregarded as unphysical. For the interested reader, in \cref{BELreg} we provide a detailed overview of this dimensional regularization of the BEL diagram.
\subsection{Two-loop obstruction}\label{2LoopObstruction}
At two-loop, introducing terms that conspire with four-point contacts
is unavoidable. At one-loop, we were able to avoid internal
contractions of $\bar{A}_\mu A^\mu$ by selecting appropriate external
states. However, at two-loop when choosing all external $A_\mu$
states, the amplitude vanishes in semi-YM theory:
\begin{equation}
\mathcal{A}^{\text{semi-YM}}_{\text{2-loop}}(A_\mu,A_\mu,A_\mu,A_\mu)=0
\end{equation}
In order to produce non-vanishing interactions, we would need to
reintroduce $D$-dimensional vertices from the diagonal group of
$G_L\times G_R$ that have the wrong weight. Doing so will necessarily permit internal
contractions of $\bar{A}_\mu A^\mu$,
\begin{equation}
\mathcal{A}^{\text{YM}}_{\text{2-loop}}(A_\mu,A_\mu,A_\mu,A_\mu)=\doubleBoxObstructionA +\doubleBoxObstructionB+\cdots 
\end{equation}
where we used white dots to indicate interaction vertices of weight\footnote{We choose the convention that the abelianized gauge fields carry weight $\mathcal{W}[\bar{A}] = -1$, and non-abelian vectors are $\mathcal{W}[\bar{A}] = +1$.} $\mathcal{W}[\bar{A}\bar{A} A] = -1$, rather than the isolated semi-YM vertex which always carries weight $\mathcal{W}[\bar{A}A A] = +1$. This immediately runs into the difficulty of introducing the
four-point contact needed for color-kinematics to be satisfied on all
internal edges. By a simple weight counting argument, one can see that including this wrong sign interactions is unavoidable at two-loop and higher.
